---
title: "EDS 223: Week 1 Discussion Section"
author: "Rachel Swick"
format: html
---
## 1. Learning Objectives
- Read in spatial objects
- Create map with single spatial object
- Create map with multiple spatial objects
- Use different types of `tmap` plotting formats (e.g. `tm_polygons()`, `tm_fill()`, `tm_dots()`, etc.)
- Adjust color palettes
- Include essential map elements (e.g. scale bar & north arrow or graticules)
- Create an interactive map
- Bonus Challenge: Reproduce map using `ggplot2` instead of `tmap`

## 2. Get Started
- Open your forked version of this repository and navigate to your version controlled project for Week 1
- Create a Quarto document
- Load all necessary packages
- Read in the spatial objects for Easter Island (Rapa Nui/Isla de Pascua):
  - `ei_points`: file contains several points on the island
  - `ei_elev`: raster with elevation data
  - `ei_borders`: polygon with the island outline
  - `ei_roads`: lines contains a road network for the island
  
```{r}
# load libraries
library(here)
library(tidyverse)
library(stars)
library(sf)
library(tmap)
```
 
```{r}
# read in the spatial objects for Easter Island
ei_points <- sf::read_sf(here::here("data", "easter_island", "ei_points.gpkg"))
volcanoes <- subset(ei_points, type == "volcano")
ei_elev <- stars::read_stars(here::here("data", "easter_island", "ei_elev.tif"))
ei_borders <- sf::read_sf(here::here("data", "easter_island", "ei_border.gpkg"))
ei_roads <- sf::read_sf(here::here("data", "easter_island", "ei_roads.gpkg"))
```
  
## 3. Your Task

1. Create a map of Easter Island
```{r}
tm_shape(ei_borders) +
  tm_polygons(col = "lightgreen",
              borders.col = "black",
              lwd = 3) +
  tm_layout(main.title = "Map of Easter Island",
            frame = FALSE)
```

2. Create a map of Easter Island and…
    …denote the island’s borders and continuous elevation
    …denote the island’s volcanoes and roads
    …play with the color palette and essential map elements
```{r}
tm_shape(ei_elev) +
  tm_raster(palette = c("tan", "brown"),
            title = "Elevation (feet)") +
tm_shape(ei_borders) +
  tm_borders() +
tm_shape(volcanoes) +
  tm_dots(size = 0.2,
          col = "orange") +
tm_shape(ei_roads) +
  tm_lines() +
tm_layout(frame = FALSE,
          title = "Map of Volcanoes on Easter Island")+
  tm_scale_bar() +
  tm_compass(position = c("left", "bottom"))
```

3. Create an interactive map of Easter Island  
```{r}
tmap_mode("view")

tm_shape(ei_elev) +
  tm_raster(palette = c("tan", "brown"),
            title = "Elevation (feet)") +
tm_shape(ei_borders) +
  tm_borders() +
tm_shape(volcanoes) +
  tm_dots(size = 0.2,
          col = "orange") +
tm_shape(ei_roads) +
  tm_lines() +
tm_layout(frame = FALSE,
          title = "Map of Volcanoes on Easter Island")+
  tm_scale_bar() +
  tm_compass(position = c("left", "bottom"))
```

  
  
  
  
  
  
  
  
  
  
  
  
  
  
  